---
title: "Learning Cpp"
author: "Dereck de Mézquita"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "figures", "cs-Cpp.html")) })
output:
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
    # code_folding: hide
params:
  rmd: cs-Cpp.Rmd
editor_options: 
  chunk_output_type: inline
---

<a download="cs-Cpp.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 7, fig.align = "center")
```

```{r cpp-engine, include=FALSE}
knitr::knit_engines$set(Cpp = function(options) {
	out_dir <- "./cpp"
	i <- max(length(list.files(out_dir, pattern = "*cpp")))
	
	execname <- stringr::str_interp('./cpp/chunk_${i}')
	filename <- stringr::str_interp('${execname}.cpp')
	
	code <- paste(options$code, collapse = '\n')
	
	system2(
		'echo', c(shQuote(code), ' > ', filename), stdout = TRUE
	)
	
	system2(
		'/usr/bin/c++', c(filename, ' -o ', execname)
	)
	
	exec <- system2(
		execname, stdout = TRUE
	)
	
	system2(
		'rm', execname
	)
	
	knitr::engine_output(options, code, exec)
})
```

```{r libraries, warning=FALSE, message=FALSE}
library("Rcpp")
library("ggplot2")
```

# Introduction to `C++`

## About this course

The point of this course is understanding the basics and some advances concepts in `C++`. For ease of use and presentation `.Rmd` documents are used to produce this course. 

A custom `C++` engine is defined for this course. Compiles chunks to an executable, runs programme and returns the stdout to the `.Rmd` document. In some cases with `Rcpp` as an interface to `C++` when useful such as for plotting etc.

Some conventions: 

1. We will always use base 0 counting; the standard in `C++`.
1. `Rcpp.h` libraries are avoided in favour of the standard `C++` headers.
1. Code will more often than not be written in a function body; allows compilation and use in `R`'s context.

Use of `Rcpp.h` libraries will be kept to a minimum and information on compilation and using `C++` outside the context of `R` will be detailed.

Let's start with something simple; demonstrate that we can indeed write and compile `C++` in this context.

Here is a function which generates the Fibonacci sequence to `n` elements.

```{Rcpp cpp-forloop-func, cache=TRUE}
#include <Rcpp.h>
using namespace Rcpp;

// 0, 1, 1, 2, 3, 5, 8, 13

// [[Rcpp::export]]
NumericVector fibonacci_cpp(int n) {
	NumericVector vec(n);
	
	vec[1] = 1;
	for(int i = 1; i < n; i++) {
		vec[i + 1] = vec[i - 1] + vec[i];
	}
	
	return vec;
}
```

```{r print-cpp-fib}
fibonacci_cpp(20)
```

## About `C++`

`C++` is a language invented by Bjarne Stroustrup in 1979. It is based on the `C` language and can be summarised as `C` with classes; object oriented programming.

`C++` is a low level, compiled, and powerful language. `C++` has become ubiquitous over the years and virtually all systems support compiling and running compiled `C++` code.

Printing "hello world" in typical `C++` would look something like this:

```{Cpp cpp-hello}
#include <iostream>

int main() {
  std::cout << "Hello World!" << std::endl;
}
```

`C++` is written in files with a `.cpp` extension. The output you see above was generated with our custom chunk engine, you can find the standalone code again in the "./cpp/" directory.

## Anatomy of a `C++` programme

A programme written in `C++` is typically structured as such:

Libraries are just code files that have been pre-written and offer some kind of utility. These typically end with a `.h` extension which stands for "header" as they are included at the top of a programme.

**Indeed** `C++` is executed in a linear fashion from first line to last.

Libraries are included at the top of a programme with the syntax: `#include <iostream>`.

Libraries can be stored locally to a project or on a system level. Libraries can be written by anyone; you could write helper functions and have it stored locally to your project.

Some libraries have become so common that they are stored in a single location for a computer. These can be included using the brackets `<name of library>` would do it. In the case that you want to include a locally stored header file you would use double quotes `"name of file"`.

A `main` function is always included in the programme and always returns an `int` (integer). This is because a success and failure to run is denoted with `0`/`1`.

A basic `C++` document would like this; note the function structure - parenthesis and curlies.

**Note**: comments are written with `//` or for multi-line between `/* comment here */`.

```{Cpp cpp-structure}
#include <iostream> // library

int main() { // declare the type returned; and main function
	// do something
}
```

In our previous example we wrote: `std::cout << "Hello world!"`. This comes from the `iostream` library, the `std` signals that we're getting the `cout` function from that library. Finally, `<<` is an operator that pases the right hand statement as an input to the left hand.

**Important: ALL* lines in `C++` must be ended with a semi-colon - ;**. All statements that is. *In some cases like the end curly of a function we do not for example.


## Workflow

As stated before `C++` is a compiled language. This means we will often follow this routine of: 

1. Write programme.
1. Save programme.
1. Compile programme.
	1. If error; read error, go back to step 1.
1. Execute programme.
	1. If error; read error, go back to step 1.

So I previously showed you what some of the code looks like. Why don't I explain to you some of the magic of how to actually compile a programme?

Indeed I've set this all up to be compiled from the code chunks in this document auto-magically with the use of a custom chunk engine. But it's super simple I promise. I'll explain myself.

So what you typically do as we said; is we write the progamme, let's write one:

```{Cpp cpp-hello-world, eval=FALSE}
#include <iostream>

int main() {
	std::cout << "Hello world." << std::endl;
}
```

Ok now save the above to a text file with the extension `.cpp`.

Now try this command in your terminal:

```{bash cpp-compile, eval=FALSE}
c++ programme_name.cpp -o programme_name
```

What are we doing there? Well we're compiling the programme from the text file, `-o` denotes the output name/location of the programme executable.

So let's try it:

```{bash cat-hello}
echo '#include <iostream>

int main() {
	std::cout << "Hello world." << std::endl;
}' > ./cpp/hello.cpp

c++ ./cpp/hello.cpp -o ./cpp/hello
```

So now we should have our programme in the cpp directory let's take a look:

```{bash run-hello}
ls ./cpp/hello*

echo ''

cat ./cpp/hello.cpp

rm ./cpp/hello
```

The text file is there and we've compiled it.

You may be thinking what is a compiled programme anyway? Why do we have to "compile"? Well a compiler is simply a translator. All it does is it takes our `C++` code and translates it to what's commonly referred to as machine code; assembly. Assembly is an extremely low level langauge that then can be directly compiled to binary which is the true language of any computer.

Indeed we are using `C++` code to talk to the comptuer. `C++` has features and concepts which allow us to more easily reason about what we want to do. 

If you want to write a bit of code that will be repeated n amount of times, in `C++` you would you something called a for loop. It looks a little something like this:

```{Rcpp cpp-for-loop, eval=FALSE}
int times = 100;
for(int i = 0; i < times; i++) {
	// do something
}
```

That is just the pure for loop. Doesn't look so intimidating does it? We will go over it more in detail later. But it's telling the computer do what's inside these brackets n times.

Let's you wanted to write assembly; baisically you would write a document and somewhere you'd tell it to JUMP back and repeat that code exactly with a new value for the iterator.

Why don't we write a loop, compile it and then try to take a look at what the assembly code looks like?

Here is a simple for loop, prints hello followed by the iterator 3 times.

```{Cpp cpp-hello-loop}
#include <iostream>

int main() {
	for(int i = 1; i <= 3; i++) {
		std::cout << "Hello " << i << std::endl;
	}
}
```

Here we echo the code into a document and then compile it. Note the `-S` argument this tells the compiler to output assembly code.


```{bash cat-hello-loop}
echo 'int main() {
	int i;
	for(i = 1; i <= 3; i++) {
		int x = i;
	}
	
	return i;
}' > ./cpp/print_loop.cpp

c++ ./cpp/print_loop.cpp -o ./cpp/print_loop.asm -S
```

So now we've saved the code to a document, and compiled it. Let's take a look. 

**Note**: we've removed the print function and `<iostream>` from the code, as this includes a whole library which we don't really need. We just want to see what for loop looks like in assembly.

```{bash print-assembly-loop}
cat ./cpp/print_loop.asm
```


In summary from the above we can see how compiling works. You write some code then you use a compiler programme which translates your code to assembly/machine code.

The command would look something like this as we used above:

```{bash compile-assembly, eval=FALSE}
c++ ./cpp/print_loop.cpp -o ./cpp/print_loop -S
```


## Data structures and data types

### Types

You've seen the `int` prefix in the code I've shown you before. Notably before the main function. As I said before all variables and functions must have their type declared. Here we'll go over some data types and data structures of `C++`.

The primitive data types, those built into `C++`, are:

1. `int`: integer numbers.
1. `double`: used for storing double precision floating point numbers/decimal numbers.
1. `float`: stores single precision floating point numbers/decimal numbers.
1. `char`: individual characters; corresponding to ASCII table value maxes out at 256.
1. `string`: a sequence of characters.
1. `wchar_t`: characters; corresponding to UNICODE table value maxes out at 65,536.
1. `bool`: a true/false or logical - 0 and 1 can also be used and cast to a `bool`.
1. `void`: no value, represents a valueless entity; typically used for functions that return nothing.

Note that you can also have modified types; these are types with a modifier:

1. `signed`: a number type which can be both positive and negative.
1. `unsinged`: a number type which can only be positive.
1. `short`: this is used for small integers in a range between -32768 and 32767.
1. `long`: used for large numbers.

### Information theory

Different types have different properties, specifically referring to memory use. On most machines (depending on the architecture) a variable takes up anywhere from 1 byte to 12 bytes of memory (called memory allocation).

What even is memory? What is information? How is information stored? When working with computers or even with other humans we all have to agree as to how we are going to represent information. One way of representing information is to count things on the digits of your hand. One finger is one, two is two etc. with this we can count to ten.

We also work with other kinds of numbers, decimal numbers; technically base 10 numbers. This refers to the fact that there are 10 symbols which are used for counting. These are 0 to 9.

$$
	0,\,1,\,2,\,3,\,4,\,5,\,6,\,7,\,8,\,9
$$

However computers don't speak the same language as us. Their language is much more simple. Computers use a base 2 system or known as binary. Indeed, they count with two symbols. For a computer a single value, either a 0 or 1, is known as a bit; this is derived from the combination of two words binary and digits.

The reason for such a simple system is because of the fact that computers are built on electronic devices called transistors. Transistors are simply gates for electricity, they either let electricity through or don't; on or off.

Here we represent counting with a single bit. We are limited to counting with two symbols on or off, or 0 to 1. Let this dot represent a "light bulb", yellow means it's on or 1.

```{r lightbulb-one-bit, echo=FALSE, fig.height=2}
par(bg = "black", mar = rep(0, 4))
plot.new()
plot.window(xlim = range(0, 2), ylim = c(1, 2.75))
points(1, 1.9, col = "yellow", pch  = 19, cex = 7)
text(1 - 0.5, 1.9, col = "yellow", labels = "ON/1", cex = 1.75)
```

You might be thinking if we only have two symbols 0 and 1, how can we count higher than 1? Well we could use more than one bit (value 0/1) in different combinations to represent different values.

Using bits in different combinations to represent numbers is something similar to what we already do with our own number system. Imagine 0 - 9; the numbers 1, 2 and 3 can be all placed next to each other to represent what we instinctively know as one-hundred twenty three.

$$
	123
$$

Note that each digit in this number represents a place of value: 1 - hundreds place, 2 tens place, and 3 the ones place. So the solution here:

$$
	123 = 100 \times 1 + 10 \times 2 + 1 \times 3
$$

The binary system is the same expect they don't use 9 symbols but 2! We have the number 10 to represent our 10 symbols (0 - 9) and the 2 the two of the computer (0 - 1). These exponents are in essence telling us how many values each place can represent; 10 ^ 2 = `r 10 ^ 2`.

$$
	 10 ^ 2, 10 ^ 1, 10 ^ 0 \\
	 2 ^ 2, 2 ^ 1, 2 ^ 0 
$$

So we humans have ones place, tens, then hundreds place. Computers with their binary systems have ones place, twos place, and fours place. Simply resolve the above expressions.

If we have 1 bit we can count from 0 to 1. How high could we count if we have three bits?

```{r lightbulb-three-bits, echo=FALSE, fig.height=5.5}
lightbulbs <- gtools::permutations(2, 3, c("yellow", "gray"), repeats.allowed = TRUE)
lightbulbs <- lightbulbs[, 3:1]

coord <- data.frame(x = unlist(as.data.frame(col(lightbulbs))), y = unlist(as.data.frame(row(lightbulbs))), row.names = NULL)

state <- 7:0
vals <- c(
	"4 * 0", "2 * 0", "1 * 0",
	"4 * 0", "2 * 0", "1 * 1",
	"4 * 0", "2 * 1", "1 * 0",
	"4 * 0", "2 * 1", "1 * 1",
	"4 * 1", "2 * 0", "1 * 0",
	"4 * 1", "2 * 0", "1 * 1",
	"4 * 1", "2 * 1", "1 * 0",
	"4 * 1", "2 * 1", "1 * 1"
)

par(bg = "black", mar = rep(0, 4))
plot.new()
plot.window(xlim = range(0.5, 3.5), ylim = c(0.25, 8.75))

points(coord$x, coord$y, col = rev(unlist(as.data.frame(lightbulbs))), pch  = 19, cex = 7)
text(coord$x[1:8] - 0.35, coord$y[1:8], col = "yellow", labels = state, cex = 1.75)
text(coord$x, coord$y, col = "black", labels = rev(vals), cex = 1)
```

The above illustration shows us what it would be like to count with 3 bits. What we did was organise our bits in all possible combinations and these can be used to represent a numerical value. Note that we start counting at 0 and thus we can only get as high as 7 if we have 3 bits. The more bits you have the higher you can count.

### Types and bytes

As you might know 1 byte is equal to 8 bits. A single bit is the smallest unit of memory. We will typically work with whole bytes as that is what is accessible to us. Working with a single individual bit is in essence impossible, a variable must have an associated address which also occupies some amount of space.

As an example note that an `int` is typically 4 bytes in size. This equates to: 4 bytes each 8 bits in size. So we have a total of 32 bits to store information in.

Okay but what in the heck is a bit? A bit as you might know is the smallest unit by which we can measure memory. It is either on or off, 0 or 1. This is commonly known as binary! 

So as we said before we have 32 0s and 1s in which to store information with an `int` data type. These can individually be in the 0 or 1 state. 

How is information even stored? For example let's think we have 0s and 1s to work with.

Typically these are the sizes of the following types:

Data Type | Size (in bytes) | Range
-- | -- | --
short int | 2 | -32,768 to 32,767
unsigned short int | 2 | 0 to 65,535
unsigned int | 4 | 0 to 4,294,967,295
int | 4 | -2,147,483,648 to 2,147,483,647
long int | 4 | -2,147,483,648 to 2,147,483,647
unsigned long int | 8 | 0 to 4,294,967,295
long long int | 8 | -(2^63) to (2^63)-1
unsigned long long int | 8 | 0 to 18,446,744,073,709,551,615
signed char | 1 | -128 to 127
unsigned char | 1 | 0 to 255
float | 4 |  
double | 8 |  
long double | 12 |  
wchar_t | 2 or 4 | 1 wide character

Let's start with some numbers. You can declare a number first, then attribute a value. The important step with anything in `C++` is delcaring the type.

```{Cpp cpp-types}
#include <iostream>

int main() {
	int yeet; // declare yeet
	
	yeet = 1; // set equal to 1
	
	std::cout << yeet << std::endl;
	
	yeet = 2; // reset to 2
	
	std::cout << yeet << std::endl;
	
	int zoot = 131; // declare and set zoot
	
	std::cout << zoot << std::endl;
	
	int xoot = yeet + zoot; // declare and set xoot as sum of yeet + zoot
	
	std::cout << xoot << std::endl;
}
```

As you can see our variable `yeet` was declared and then set to equal 1, then subsequently equal to two and printed again.

**Note**: adding and arithmetic functions are as natural as +/-. Let's go over some maths yes?

## Arithmetic operators

In `C++` we have these as the following primitive arithmetic operators. Primitive because they are built into the language. You'll notice when you start needing more complex things such as n to the power of x you'll need some libraries.

- `+` addition.
- `-` subtraction.
- `*` multiplication.
- `/` division.
- `%` modulo; gets the remainder after the division.

```{Cpp operators}
#include <iostream>

int main() {
	double score = 0;
	
	score = score + 20;
	std::cout << score << std::endl;
	
	score = score - 2;
	std::cout << score << std::endl;
	
	score = score * 3;
	std::cout << score << std::endl;
	
	score = score / 2;
	std::cout << score << std::endl;
	
	score = (int) score % 3;
	std::cout << score << std::endl;
}
```


Notice in the last operation I put `(int)` before taking modulo 3, this is something called type casing and we will learn more about it in the future.


## Project: quadratic maths

Let's put together what we've learned and make a programme that can solve the quaradtic formula. We'll also visualise what we do with some `R` code.

What even is the quadratic formula? Well a quadratic formula is a mathematical formula that allows you to find the solutions to a quadratic equation... 

I know, never define a word by using the word in the definition! Stick with me. A quadratic (meaning square or 4 in latin) equation is an equation that fits the form of:

$$
	ax ^ 2 + bx + c = 0
$$

1. `x` represents a variable.
1. `a, b, c` represent known numbers; `a != 0`.

What's going on in this equation to put things simply; you input a value for `x` and you get `y` back. You can then plot these!

Shifting the values of `a, b, c` is what defines different quadratic functions.

Note that if a is equal to zero then the equation simply becomes linear: 

$$
	bx + c = 0
$$

Let's try it out and visualise. Here we define a function to calculate quadratic coordinates.

In order to get the values back for plotting we'll use `R`'s flavour of `C++`. We'll use `Rcpp`'s header file. Some other changes to note is no `main()` function, this is because the code is compiled by `Rcpp` the imported as a function to `R`.

Note the `using namespace Rcpp;` line; this means we can directly use the library without having to append everything with `Rcpp::`function_name. You can do the same thing with `std`.

Let's dissect the following code. This is 99% `C++` with a little bit of `Rcpp` in there just for getting back the function in here.

```{Rcpp quad-return-y}
#include <Rcpp.h>
#include <vector>

// [[Rcpp::export]]
std::vector<double> quadraticReturnY(std::vector<double> x, double a, double b, double c) {
	int n = x.size();
	
	std::vector<double> y(n);
	
	for(int i = 0; i < n; i++) {
		y[i] = (a * (x[i] * x[i])) + (b * x[i]) + c;
	}
	
	return y;
}
```

Here you see `<vector>`, what is a vector? A `vector` is a sort of super array defined in the standard library. It has some nice methods attached to it so we can get it's size easily like shown above for example: `x.size();`

Note that we are creating a `vector` with the following syntax: `std::vector<double>`. This is saying get the `vector` class from the standard library, and we're putting numbers inside this `vector` of type double.

What we're saying here is input to this function a `vector`, and some other variables. Do some calculations and return the `vector` of results y.

Let's then calculate the coordinates for a quadratic function of the following parameters:

1. `a = 1`
1. `b = 0`
1. `c = 0`

```{r run-quady}
a <- 1
b <- 0
c <- 0

x <- -20:20
y <- quadraticReturnY(x, a, b, c)

df <- data.frame(x = x, y = y)

head(df, 5)
```

```{r plot-classic-quadratic}
if(df$y[which.max(abs(df$y))] > 0) {
	lim <- c(-15, NA)
} else {
	lim <- c(NA, 15)
}

ggplot2::ggplot(df, ggplot2::aes(x = x, y = y)) +
	ggplot2::geom_line() + 
	ggplot2::scale_y_continuous(limits = lim) +
	ggplot2::geom_hline(yintercept = c, linetype = "dashed", colour = "goldenrod", size = 1) +
	ggplot2::labs(title = "A classic quadratic plot", subtitle = stringr::str_interp('a = ${a}, b = ${b}, c = ${c}, x = [${min(x)} ... ${min(x)}]'))
```

Notice how if we shift the values we move the plot around:

```{r plot-shifted-quadratic}
a <- 1.4
b <- 40
c <- -4

x <- -30:20
y <- quadraticReturnY(x, a, b, c)

df <- data.frame(x = x, y = y)

ggplot2::ggplot(df, ggplot2::aes(x, y)) +
	ggplot2::geom_line() + 
	ggplot2::geom_hline(yintercept = -4, linetype = "dashed", colour = "goldenrod", size = 1) +
	ggplot2::labs(title = "A shifted quadratic plot", subtitle = stringr::str_interp('a = ${a}, b = ${b}, c = ${c}, x = [${min(x)} ... ${min(x)}]'))
```

So in the examples above we're asking the formula to give us back the `y` values for an array of `x`s.

Notice how the parabola at some point crosses the x axis. At this point y really does equal 0. This is called a solution, or a root to this equation. The equation is defined by the a, b and c constants in the context of the quadratic formula shown before.

There exists two roots, as a parabola has two arms, these are two solutions or two points where the arms intersect the x-axis. These points can be found by plugging these same constants into this formula.

Note if your parabola is shifted above the x-axis it's possible you have no real solutions.

Let's implement this formula for calculating a solution to the above shown parabola.

$$
	x = \frac{-b \pm \sqrt{b ^ 2 - 4ac}}{2a}
$$

```{Rcpp solve-quadratic}
#include <Rcpp.h>
#include <math.h>
using namespace Rcpp;

// [[Rcpp::export]]
std::vector<double> quadraticRoots(double a, double b, double c) {
	
	std::vector<double> x(2);
	
	x[0] = (-b + std::sqrt((b * b) - (4 * a * c))) / (2 * a);
	x[1] = (-b - std::sqrt((b * b) -( 4 * a * c))) / (2 * a);
	
	return x;
}
```

Wonderful as we can see the roots or solutions to this quadratic formula are these! Note that if you were to plot these as points on the graph where these correspond to the x value (if y = 0) then you would find that these are indeed the points where the parabola's arms cross the x-axis.

```{r print-quad-solution}
quadraticRoots(a, b, c)
```

Here's the same plot again, with the points we just calculated as red circles! Amazing!

```{r plot-quadratic-solutions}
df2 <- data.frame(x = quadraticRoots(a, b, c), y = rep(0))

ggplot2::ggplot() +
	ggplot2::geom_line(data = df, ggplot2::aes(x, y)) +
	ggplot2::geom_point(data = df2, ggplot2::aes(x, y), colour = "red", size = 5, alpha = 0.5) +
	ggplot2::geom_hline(yintercept = -4, linetype = "dashed", colour = "goldenrod", size = 1) +
	ggplot2::labs(title = "A shifted quadratic plot", subtitle = stringr::str_interp('a = ${a}, b = ${b}, c = ${c}, x = [${min(x)} ... ${min(x)}]'))
```


## Conditionals and logic

In `C++` conditonals and control flow is pretty straight forward as in most languages:

```{Cpp}
#include <iostream>

int main() {
	bool yeet = true;
	std::string mom_location = "home";
	int people = 9;
	
	if(yeet & (mom_location != "home")) {
		std::cout << "Mother's home no yeeting!" << std::endl;
	} else if(((bool) people % 3) & yeet) {
		std::cout << "We have enough yeets for all and mother is not home! YEET!" << std::endl;
	} else {
		std::cout << "We don't have enough yeets, but we have some!" << std::endl;
	}
}
```

Above we demonstrate a number of concepts. Let's enumerate them here:

- `==` a comparator, is left side equal to right side.
- `!=` not equal to.
- `>` greater than.
- `<` less than.
- `>=` greater than or equal to.
- `<=` less than or equal to.

Notice how we're catching if the first condition is not met then the second is checked, finally the last statement is executed as a catch all.

If we start having a large number of conditions things can get convoluted. A `switch` statement is suggested.

Note a switch must always evaluate to a integral type: `int`, `char`, `short`, `long`, `long long`, `enum`.

The case keyword checks if the expression fed matches the value after. Break tells the computer to exit the block and not run the other cases. At the end there is a default case as a catch all.


```{Cpp}
#include <iostream>
#include <random>

int main() {
	int bedrooms = 9;
	
	switch(bedrooms) {
		case 1 :
			std::cout << "The house is cheap" << std::endl;
			break;
		case 2 :
			std::cout << "Starter home" << std::endl;
			break;
		case 3 :
			std::cout << "Mid sized house for a family." << std::endl;
			break;
		case 4 :
			std::cout << "Okay big house!" << std::endl;
			break;
		default :
			std::cout << "Why do you need so many rooms!?" << std::endl;
			break;
	}
}
```

## Logical operators

Above you might have noticed I used `&` inside one of the if statements. This is a conditional operator. Conditional operators combine booleans and output a single condition. In essense these are different kinds of logic gates. 

- true + true = true
- true + false = false
- false + true = false

And so on.

- `&&` and.
- `||` or.
- `!` not.

```{Cpp}
#include <iostream>

int main() {
	std::cout << "1 < 2 && 2 > 3: " << (1 < 2 && 2 > 3) << std::endl;
	std::cout << "1 < 2 || 2 > 3: " << (1 < 2 || 2 > 3) << std::endl;
}
```

## Project: leap year?

Let's write a small programme that checks if a given year is a leap year.

```{Rcpp}
#include <Rcpp.h>
#include <vector>

// [[Rcpp::export]]
std::vector<double> quadraticReturnY(std::vector<double> x, double a, double b, double c) {
	int n = x.size();
	
	std::vector<double> y(n);
	
	for(int i = 0; i < n; i++) {
		y[i] = (a * (x[i] * x[i])) + (b * x[i]) + c;
	}
	
	return y;
}
```

```{Rcpp}
#include <Rcpp.h>
#include <iostream>
#include <vector>

// [[Rcpp::export]]
void leapYear(std::vector<int> year) {
	std::vector<bool> leap(year.size());
	for(int i = 0; i < year.size(); i++) {
		if(std::to_string(year[i]).length() != 4) {
			std::cout << "Number is not length 4." << std::endl;
			throw;
		}
	}
}
```


